---
title: "Red Wines Analysis"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Red wine analysis
first let's read the data and load some usfull libiraries
```{r}
Mydata=read.csv("C:/Users/ahmed/Downloads/wineQualityReds.csv")
library('ggplot2')
library('ggthemes')
library('gridExtra')
library('dplyr')
library('psych')
```
ok, after that first thing i love to do is to look at the distribution of every variable 
```{r}
d1<-ggplot(data = Mydata,aes(quality))+geom_bar()
d2<-ggplot(data = Mydata,aes(fixed.acidity))+geom_histogram(bins = 15)
d3<-ggplot(data = Mydata,aes(volatile.acidity))+geom_histogram(bins = 20)
d4<-ggplot(data = Mydata,aes(citric.acid))+geom_histogram(bins = 20)
grid.arrange(d1,d2,d3,d4)
d5<-ggplot(data = Mydata,aes(residual.sugar))+geom_histogram(bins = 20)
d6<-ggplot(data = Mydata,aes(chlorides))+geom_histogram(bins = 50)
d7<-ggplot(data = Mydata,aes(total.sulfur.dioxide))+geom_histogram(bins = 50)
d8<-ggplot(data = Mydata,aes(free.sulfur.dioxide))+geom_histogram(bins = 15)
grid.arrange(d5,d6,d7,d8)
d9<-ggplot(data = Mydata,aes(sulphates))+geom_histogram(bins = 15)
d10<-ggplot(data = Mydata,aes(pH))+geom_histogram(bins = 15)
d11<-ggplot(data = Mydata,aes(density))+geom_histogram(bins = 15)
d12<-ggplot(data = Mydata,aes(alcohol))+geom_histogram(bins = 15)
grid.arrange(d9,d10,d11,d12)
```

*Quality has most values concentrated in the categories 5, 6 and 7. Only a small proportion is in the categories [3, 4] and [8, 9] and none in the categories [1, 2] and 10.
*Residual sugar has a positively skewed distribution.
*Alcohol has an irregular shaped distribution.
let's use some descriptive statistics that might help.
```{r}
quality_min <- min(Mydata$quality)
quality_max <- max(Mydata$quality)
quality_mean <- mean(Mydata$quality)
quality_median <- median(Mydata$quality)
```
We can see that, although the range of posible scores is from 0 to 10, in our dataset the minimum score is 3 and the maximum is 8. The mean is 5.6360225 and the median is 6, very close to each other.

What about the dispersion around the mean?
```{r}
reds_quality_sd <- sd(Mydata$quality)
```
With a standard deviation of 0.8075694, our distribution of quality scores is not very disperse.

Let have a look at how correlated our different variables are.
```{r}
cor(x=Mydata[,2:12], y=Mydata$quality)
```
we see here that some variables have high correlated values like 
*Alcohol (+++)
*Volatile acidity (—)
*Citric acid (++)
I will concentrate in the this three variables that show stronger correlation with quality.
First of all, a summary of the differnet mean values.
```{r}
data_quality<-group_by(Mydata,quality)
Mydata.quality_means<-summarise(data_quality,fixed_acid_mean=mean(fixed.acidity),
          volatile_acid_mean=mean(volatile.acidity),
          citric_acid_mean=mean(citric.acid),
          sugar_mean=mean(residual.sugar),
          salt_mean=mean(chlorides),
          sulfur_mean=mean(total.sulfur.dioxide),
          sulphates_mean=mean(sulphates),
          density_mean=mean(density),
          alcohol_mean=mean(alcohol)
          )
summary(Mydata$alcohol)
alcohol_mean <- mean(Mydata$alcohol)
alcohol_median <- median(Mydata$alcohol)
Mydata.quality_means[,c('quality','alcohol_mean')]
```
#### Alcohol
let's visualize how the alcohol mean affect wine quality
```{r}
ggplot(aes(x=alcohol_mean,y=quality),data = Mydata.quality_means)+geom_line()+geom_point(aes(color=quality))
```
from this two plots we can see that with more alcohol level the wine quality increase
lets also see alcohol by quality level, together with the median for the whole distribution and the mean for the quality scores.
```{r}
ggplot(data=Mydata, aes(x=quality, y=alcohol)) +
  geom_boxplot(aes(group=cut_width(quality,1))) +
  geom_hline(show.legend =T, yintercept=alcohol_mean, linetype='longdash', alpha=.5, color='blue') +
  geom_vline(xintercept =quality_mean, linetype='longdash', color='blue', alpha=.5) +
  xlab("Wine Quality") +
  ylab("Alcohol")
```
We can see how wine scores that are beyond the mean quality value of 5.6360225, also tend to show values beyond the mean alcohol value of 10.4229831 there is some tendency.

#### Citric acid
let’s perform a similar analysis for citric acid
```{r}
summary(Mydata$citric.acid)
citric.acid_mean <- mean(Mydata$citric.acid)
citric.acid_median <- median(Mydata$citric.acid)
Mydata.quality_means[,c('quality','citric_acid_mean')]
```
The mean value for the citric acid is 0.2709756 with a standard deviation of 0.1948011. As a note, the maximum value for citric acid is 1. That is 3.7424031 standard deviations beyond the mean!probably an outlier.
```{r}
ggplot(aes(x=citric_acid_mean,y=quality),data = Mydata.quality_means)+geom_line()+geom_point(aes(color=quality))
ggplot(data=Mydata, aes(x=quality, y=citric.acid)) +
  geom_boxplot(aes(group=cut_width(quality,1))) +
  geom_hline(show.legend =T, yintercept=citric.acid_mean, linetype='longdash', alpha=.5, color='blue') +
  geom_vline(xintercept =quality_mean, linetype='longdash', color='blue', alpha=.5) +
  xlab("Wine Quality") +
  ylab("citric acid")
```

we can see here that the more citric acid the more good wine.

#### volatile acidity

lets make same summarize for volatile acidity but i will explore it in diffirent way since this variable have negativ correlation fro quality so i will categorize wines <5 as bad and >5 as good and see levels of volatile acidity for each category.
```{r}
summary(Mydata$volatile.acidity)
volatile.acidity_mean <- mean(Mydata$volatile.acidity)
volatile.acidity_median <- median(Mydata$volatile.acidity)
Mydata.quality_means[,c('quality','volatile_acid_mean')]
Mydata <- transform(Mydata, is.bad = quality<5)
bad.reds <- table(Mydata$is.bad)
bad.reds
```
now lets visualize it 
```{r}
ggplot(aes(x=volatile_acid_mean,y=quality),data = Mydata.quality_means)+geom_line()+geom_point()
ggplot(data=Mydata, aes(x=is.bad, y=volatile.acidity)) +
  geom_boxplot() +
  geom_hline(show.legend =T, yintercept=volatile.acidity_mean, linetype='longdash', alpha=.5, color='red')+
  scale_x_discrete(labels=c('good','bad'))+
  xlab("Wine quality") +
  ylab("Volatile Acidity")
```

actually let's see same plot for the three variables with high corr value
```{r}
ggplot(data=Mydata, aes(x=is.bad, y=citric.acid)) +
  geom_boxplot() +
  geom_hline(show.legend =T, yintercept=citric.acid_mean, linetype='longdash', alpha=.5, color='red')+
  scale_x_discrete(labels=c('good','bad'))+
  xlab("Wine quality") +
  ylab("citric Acidity")

ggplot(data=Mydata, aes(x=is.bad, y=alcohol)) +
  geom_boxplot() +
  geom_hline(show.legend =T, yintercept=alcohol_mean, linetype='longdash', alpha=.5, color='red')+
  scale_x_discrete(labels=c('good','bad'))+
  xlab("Wine quality") +
  ylab("alcohol")
```

in case of alcohol it seems that it dosn't affect the bad wines but with citric acid we see that at certain level of the acid the wine quality increase.

## summary&reflection

About good wines. It seems clear that they have certain levels of alcohol and citric acid, and lack of defects. We already saw charts and distribution in previous sections showing a positive correlation between these variable and quality.
What about defects? We already saw thet volatile acidity is very determinant for faulty wines. We also saw that alcohol is not. Both faulty and correct wines have similar distributions of levels of alcohol. And we finally saw that a high level of citric acid is associated with less faulty wines.